Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''

### 2023-11-19 alloscore_manuscript/supplement.rnw
more on properness...

Opportunity loss \eqref{eqn:opp-loss} rearranges to 
\begin{align}
l_o(x,Y) &= C(x - Y)_{+} + (L-C)(Y-x)_{+} \\
&= L(1-\alpha)(x - Y)_{+} + L\alpha(Y-x)_{+},
\end{align}
a form in which it is often called \emph{pinball} loss, despite its graph being an unlikely pinball trajectory for $\alpha \neq 1/2$.

@@

Expanding on the decision theoretic perspective sketched in Section \ref{sec:quantiles_shortage}, we can view a loss function 
as a general tool for formalizing a decision problem that assigns numerical value to the \emph{result} of taking an \emph{action} $x$
in preparation for an \emph{outcome} $y$.  A \emph{scoring rule} $S$ is a loss function where the action 
is a probabilistic forecast $F$ of the outcome $y$ (or the statement of $F$ by a forecaster). Just as in 
Section \ref{sec:quantiles_shortage}, given an action $F$, $S$ transforms a random outcome variable $Y$ into a random loss $S(F,Y)$. 
We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$, and the process of evaluating $S(F,y_i)$ for
a data set $\mathcal{Y}=\{y_i\}$ as \emph{scoring} $F$ against $\mathcal{Y}$.

Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their own 
(simulated) outcome data, against which they can be scored using $S$. $S$ therefore commits a probabilistic forecast $F$
to the ``self-assessment'' $\Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined by sampling from $F$, as well
to an assessment $\Ex [S(G, Y^F)]$ of any alternative forecast $G$.
A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other 
forecast $G$ as being better than $F$ itself, that is, that
\begin{align}
\Ex [S(F, Y^F)] \leq \Ex [S(G, Y^F)]
\end{align} 
for any $F,G$. Otherwise, the optimal decision for some forecaster would be to state a forecast $G$ other than the forecast $F$
which they believe describes the outcome $Y$.
A scoring rule meeting this criterion is called \emph{proper}.  The inequality can also be written 
as $\Ex_F [S(F, Y)] \leq \Ex_F [S(G, Y)]$ where the subscript specifies the distribution of $Y$.  
$S$ is \emph{strictly proper} when this inequality is sharp, in which case the \emph{only} optimal decision for a forecaster is to state
the forecast they believe to be true.

@@

But given an auxiliary decision problem $\mathcal{D}_{\mathrm{Aux}} =\{\mathcal{X}, \mathcal{Y}, l(x,y)\}$, such as the
allocation problem from the main text or the cost-loss problem from Section \ref{sec:quantiles_shortage}, we can
produce a scoring rule $S_l$ that is automatically proper using the Bayes act formalism introduced in Section \ref
{sec:methods.detailed.specific_allocation}. This works by defining $S_l(F, y)$ as the value $l(x^F,y)$ already given by  
$\mathcal{D}_{\mathrm{Aux}}$ for the Bayes act $x^F$; that is, the action which \emph{by design} will be assessed via
the forecast $F$ to have an expected loss $\Ex[l(x^F,Y^F)]$ as low as possible.

@@

>>>>

Opportunity loss \eqref{eqn:opp-loss} rearranges to 
\begin{align}
l_o(x,Y) &= C(x - Y)_{+} + (L-C)(Y-x)_{+} \\
&= L(1-\alpha)(x - Y)_{+} + L\alpha(Y-x)_{+} \\
&= L(\alpha - \mathbf{1}\{Y < x\})(Y-x),
\end{align}
a form in which it is often called \emph{pinball} loss, despite its graph being an unlikely pinball trajectory for $\alpha \neq 1/2$.

@@

Expanding on the decision theoretic perspective sketched in Section \ref{sec:quantiles_shortage}, we can view a loss function 
as a general tool for formalizing a decision problem that assigns numerical value to the \emph{result} of taking an \emph{action} $x$
in preparation for an \emph{outcome} $y$.  A \emph{scoring rule} $S$ is a loss function where the action 
is a probabilistic forecast $F$ of the outcome $y$ (or the statement of $F$ by a forecaster). Just as in 
Section \ref{sec:quantiles_shortage}, given an action $F$, $S$ transforms a random outcome variable $Y$ into a random loss $S(F,Y)$. 
We refer to the realized loss $S(F,y)$ as the \emph{score} of $F$ at $y$, and the process of evaluating $S(F,y_i)$ for
a data set $\mathcal{Y}=\{y_i\}$ as \emph{scoring} $F$ against $\mathcal{Y}$. Perhaps the most fundamental example of a scoring rule
from a statistical perspective is the logarithmic score $S_{\log}(F,y) = -\log f(y)$ (where $f(y)$ is the density or the mass of $F$ at $y$
if it exists and otherwise the mass), that is, the negative log-likelihood of $F$ interpreted as a parameter for the singleton 
data set $\{y\}$. This frames a maximum likelihood estimate as the solution to a decision problem following the expected loss
minimization principle introduced in Section \ref{sec:quantiles_shortage}.

Decision theoretically, probabilistic forecasts are a unique kind of action in that they can be used to generate their
own(simulated) outcome data, against which they can be scored using $S$. $S$ therefore commits a probabilistic forecast
$F$ to the ``self-assessment'' $\Ex [S(F, Y^F)]$, where $Y^F \sim F$ is the random variable defined by sampling from
$F$, as well to an assessment $\Ex [S(G, Y^F)]$ of any alternative forecast $G$. For $S_{\log}$ this self-assessment is
the Shannon entropy $H(F) = -\int \log(f(x))f(x)dx$ of $F$, and adding to $H(F)$ the Kullback-Leibler (KL) divergence $D_
{KL}(F,G) = -\int \log(g(x)/f(x))f(x)dx$ gives $F$'s assessment $\Ex [S_{\log}(G, Y^F)]$ of $G$. (That is, $\Ex [S_
{\log}(G, Y^F)] = H(F) + D_{KL}(F,G)$.) The KL divergence is the degree to which $F$ perceives $G$ as divergent from
being able to minimize expected loss in this particular forecaster decision problem (i.e. maximize expected log-likelihood).

A natural consistency criterion for $S$ is that it does not commit $F$ to assessing any other 
forecast $G$ as being better than $F$ itself, that is, that
\begin{align}
\Ex [S(F, Y^F)] \leq \Ex [S(G, Y^F)] \label{eqn:prop_ineq}
\end{align} 
for any $F,G$. Otherwise, the optimal decision for some forecaster would be to state a forecast $G$ other than the forecast $F$
which they believe describes the outcome $Y$.
A scoring rule meeting this criterion is called \emph{proper}.   The inequality can also be written 
as $\Ex_F [S(F, Y)] \leq \Ex_F [S(G, Y)]$ where the subscript specifies the distribution of $Y$.  
$S$ is \emph{strictly proper} when this inequality is sharp, in which case the \emph{only} optimal decision for a forecaster is to state
the forecast they believe to be true. The logarithmic scoring rule $S_{\log}$, for example, is strictly proper due to the 
positivity of KL divergence.

@@

But given an auxiliary decision problem $\mathcal{D}_{\mathrm{Aux}} = \{\mathcal{X}, \mathcal{Y}, l(x,y)\}$, such as the
allocation problem from the main text or the cost-loss problem from Section \ref{sec:quantiles_shortage}, we can
produce a scoring rule $S_l$ that is automatically proper using the Bayes act formalism introduced in Section \ref
{sec:methods.detailed.specific_allocation}. This works by defining $S_l(F, y)$ as the value $l(x^F,y)$ already given by  
$\mathcal{D}_{\mathrm{Aux}}$ for the Bayes act $x^F$; that is, the action which \emph{by design} will be assessed via
the forecast $F$ to have an expected loss $\Ex[l(x^F,Y^F)]$ as low as possible.

@@

The study of \emph{elicitability} also concerns this construction (see, e.g., \cite{gneiting2011making}), but from a
different vantage point than ours. The question there is whether for a given map (called a \emph{functional} in
statistics and a \emph{property} in computer science) $T(F)=x \in \mathcal{X}$ of distributions into the auxiliary
action space, there exists a loss function $l$ such that $T(F) = x^F$ is the Bayes act for $l$ and the associated
scoring rule $S_l(F,x)$ is strictly proper. Such an $l$ is said to \emph{elicit} $T$.  Our focus is rather on how
forecast evaluation proceeds via the Bayes act construction for a given loss function of subject matter interest.

\bibliography{allocation}
